---
title: "Cleaning NSFG Data"
subtitle: 'Notebook to clean NSFG 2017-2019 data'
author: "Seyram A. Butame"
date: "April 2022"
output:
  html_notebook:
    toc: yes
    number_sections: yes
    toc_float: yes
    smooth_scroll: true
---

# INTRODUCTION

Some months ago, a client wanted some charts to help visualize the risk factors that impact youth HIV. The National Survey of Family Growth, one of the nationally-representative health surveys conducted by the CDC, is a good source of information to meet the client’s needs. The data is publicly available via the division of [National Center for Health Statistics](https://www.cdc.gov/nchs/nsfg/index.htm). 

To create these visuals, I have chosen to use the most recent iteration of the NSFG (at the time of this document), which is 2017 - 2019. The data is available in several formats, including SAS, Stata, and SPSS. I prefer Stata, so I downloaded the data in the Stata format. It is those Stata formatted data files that I work on below. I believe it is possible to import the raw data, formatted as text files (saved as *.dat*). The advantage of using the pre-formatted data files is that they contain the appropriate variable and value labels.

# R PACKAGES

```{r, warning=FALSE, message=FALSE, include=FALSE}

library(rio)
library(tidyverse)
library(readxl)
library(janitor)
library(expss)
library(lubridate)
library(ggplot2)
library(kableExtra)


```

# IMPORT DATA {.tabset}


```{r, warning=FALSE, message=FALSE}

a1 <- import("~/Documents/GitHub/NSFGViz/data/20172019male.dta")

a2 <- import("~/Documents/GitHub/NSFGViz/data/variables.csv") ## List of relevant variables

a3 <- import("~/Documents/GitHub/NSFGViz/data/20172019fem.dta")

```

## Limit DataFrame
In the imported raw data, there are too many variables. It represents the full NSFG survey. We need to limit the dataframe so we are lonly looking at specific variables. Reference **Appendix A** below, for a list of all the relevant variables. 

```{r, warning=FALSE, message=FALSE}

a4 <- a1 %>%
  select(caseid,givoralm,getoralm,oralcondm,ANALSEX2,ANALCONDM1, ANALSEX3, ANALCONDM2, malesex, malprtage,samesexany,malprthrace,SAMORAL12, RECEPANAL12, INSERANAL12, SAMESEX1, msamerel, msmnonmon, MALSHT12,JOHN2FRQ,PROS2FRQ,HIVMAL12,MSMSORT12,cndlsmal,samlifenum,samyearnum, condalls, mflastp, whycond, dateapp, attract,ORIENT_A,ORIENT_B, confconc, timalon, RISKCHEK1,RISKCHEK2, RISKCHEK3, RISKCHEK4,rectdouch,STDTST12, STDSITE12, STDTRT12,gon, chlam,herpes, genwarts, syphilis,evrinject,evrshare,VAGNUM12,ORALNUM12,ANALNUM12,nonmonog,NNONMONOG1,NNONMONOG2,NNONMONOG3,FEMSHT12,johnfreq,prostfrq,HIVFEM12,opplifenum,oppyearnum,femtouch,vagsex,agevagr,condvag,COND1BRK,COND1OFF,whycondl,getoralf,condfell,givoralf,timing,analsex,condanal,condsexl,anyoral,oppsexany,oppsexgen,SMK100,agesmk,SMOKE12,smkstop,DRINK12,UNIT30D,DRINK30D,drinkday,BINGE30,drnkmost,BINGE12,POT12,COC12,CRACK12,CRYSTMTH12,INJECT12,samesex,donblood,hivtest,nohivtst,WHENHIV_M,WHENHIV_Y,hivtstyr,hivresult,whynoget,plchiv,RHHIVT1,RHHIVT21,RHHIVT22,RHHIVT23,RHHIVT24,RHHIVT25,hivtst,prephiv,talkdoct,AIDSTALK01,AIDSTALK02,AIDSTALK03,AIDSTALK04,AIDSTALK05,AIDSTALK06,AIDSTALK07,AIDSTALK08,AIDSTALK09,AIDSTALK10,AIDSTALK11,cmhivtst,SXMON12,P12MOCONO,P12MOCON,sexfreq,confreq,sexstat,eversex,sxmtonce,ynosex,TALKPAR1,TALKPAR2,TALKPAR3,TALKPAR4,TALKPAR5,TALKPAR6,TALKPAR7,sedno,SEDNOLC1,SEDNOLC2,SEDNOLC3,SEDNOLC4,sednog,sednosx,sedbc,SEDBCLC1,SEDBCLC2,SEDBCLC3,SEDBCLC4,sedbcg,sedbcsx,sedwhbc,SEDWHLC1,SEDWHLC2,SEDWHLC3,SEDWHLC4,sedwhbcg,sedwbcsx,sedcond,SEDCONLC1,SEDCONLC2,SEDCONLC3,SEDCONLC4,sedcondg,sedconsx,sedstd,SEDSTDLC1,SEDSTDLC2,SEDSTDLC3,SEDSTDLC4,sedstdg,sedstdsx,sedhiv,SEDHIVLC1,SEDHIVLC2,SEDHIVLC3,SEDHIVLC4,sedhivg,sedhivsx,sedabst,SEDABLC1,SEDABLC2,SEDABLC3,SEDABLC4,sedabstg,sedabssx,rhadsex,ager,fmarital,rmarital,hieduc,hispanic,race,WGT2017_2019,secu,sest)

a5<-a3 %>% 
  select(ager,fmarital,rmarital,hieduc,hispanic,race,caseid,sedhiv,SEDHIVLC1,SEDHIVLC2,SEDHIVLC3,SEDHIVLC4,sedhivg,sedhivsx,hivtest,nohivtst,WHENHIV_M,WHENHIV_Y,cmhivtst,hivtstyr,hivresult,whynoget,plchiv,RHHIVT1,RHHIVT21,RHHIVT22,hivtst,prephiv,talkdoct,AIDSTALK01,AIDSTALK02,AIDSTALK03,AIDSTALK04,AIDSTALK05,AIDSTALK06,AIDSTALK07,AIDSTALK08,AIDSTALK09,AIDSTALK10,AIDSTALK11,vagsex,agevagr,condvag,whycondl,getoralm,givoralm,condfell,timing,analsex,condanal,condsexl,anyoral,oppsexany,oppsexgen,givoralf,getoralf,femsex,SAMESEX1,fsamerel,samesexany,samlifenum,samyearnum,mflastp,dateapp,attract,ORIENT_A,ORIENT_B,confconc,timalon,RISKCHEK1,RISKCHEK2,RISKCHEK3,RISKCHEK4,chlamtst,STDOTHR12,STDTRT12,gon,chlam,herpes,genwarts,syphilis,evrinject,evrshare,TALKPAR1,TALKPAR2,TALKPAR3,TALKPAR4,TALKPAR5,TALKPAR6,TALKPAR7,sedno,SEDNOLC1,SEDNOLC2,SEDNOLC3,SEDNOLC4,sednog,sednosx,sedbc,SEDBCLC1,SEDBCLC2,SEDBCLC3,SEDBCLC4,sedbcg,sedbcsx,sedwhbc,SEDWHLC1,SEDWHLC2,SEDWHLC3,SEDWHLC4,sedwhbcg,sedwbcsx,sedcond,SEDCONLC1,SEDCONLC2,SEDCONLC3,SEDCONLC4,sedcondg,sedconsx,sedstd,SEDSTDLC1,SEDSTDLC2,SEDSTDLC3,SEDSTDLC4,sedstdg,sedstdsx,sedhiv,SEDHIVLC1,SEDHIVLC2,SEDHIVLC3,SEDHIVLC4,sedhivg,sedhivsx,sedabst,SEDABLC1,SEDABLC2,SEDABLC3,SEDABLC4,sedabstg,sedabssx,onown,ONOWN18,eversex,rhadsex,ynosex,grfstsx,sxmtonce,methdiss,whencond,REASCOND01,REASCOND02,REASCOND03,REASCOND04,REASCOND05,REASCOND06,REASCOND07,vagsex,agevagr,condvag,whycondl,getoralm,givoralm,condfell,timing,analsex,condanal,condsexl,anyoral,oppsexany,oppsexgen,VAGNUM12,ORALNUM12,ANALNUM12,bisexprt,nonmonog,NNONMONOG1,NNONMONOG2,NNONMONOG3,MALSHT12,prostfrq,johnfreq,HIVMAL12,opplifenum,oppyearnum,nobedyr,stayrel,evsuspen,gradsusp,SMK100,agesmk,SMOKE12,smkstop,DRINK12,UNIT30D,DRINK30D,drinkday,BINGE30,drnkmost,BINGE12,POT12,COC12,CRACK12,CRYSTMTH12,INJECT12,WGT2017_2019,secu,sest)

```


# APPENDIX

## Appendix-A.

**Relevant Variables**

We use the kableExtra package to create beautiful tables. See the following for instructions:

 *    [kableExtra Cran R-project](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#Bootstrap_theme)

```{r}
appdx_a <- kable(a2, col.names = c("Variable Name", "Variable Label","NSFG Data"), 
      align = c("l", "l", "l")) %>%
  kable_paper("hover", full_width = F)

## Print table
appdx_a
```

